<!-- 
This is a child document for presentation the  effects of input require filtering on reference data
to be used with report-input-filt.Rmd

 -->



```{r include=FALSE,cache=FALSE}

library(knitr)
library(bookdown)


knitr::opts_chunk$set(echo = FALSE, 
                      collapse = TRUE, 
                      warning = FALSE, 
                      message = FALSE)

```

```{r}
require(ggplot2)
require(tibble)
require(tidyr)
require(dplyr)
require(reshape2)
require(viridis)
require(parallel)
require(gridExtra)
require(kableExtra)
```




`r {paste("# Effect of Filtering on Reference Data", refdata.pref, sep=" ")}`



## Read and RSL counts

Read counts per RSL-sgRNA histograms in the reference data set filtered using Input data prefiltered under different read count cut offs (see section above) are presented on Figures below.

Full scales are to present data range, and limited scales (X scale after log10 transformation) to focus on data rich region. 
Limits to Y scale are set to include maximal read count in range 1-50 under all filtering cut offs.


### Histogram plots - full data range




```{r include=TRUE,cache=FALSE}

#my.cutoffs=c(1,2,3,4,5,7,10)

my.cutoffs=c(0,1,2,3,4,5,7,10)


hist.reads_per_RSLsgRNA <- list()
hist.reads_per_RSLsgRNA2 <- list()
hist.reads_per_RSLsgRNA2.log <- list()


hist.RSL_per_sgRNA <- list()

box.RSL_per_sgRNA <- list()

detected_sgRNAs <- list()

RSL_data.list <- list()



for (i in my.cutoffs){
  filt.reflib.path=file.path(inputdata.path,paste(inputlib.name,i,refdata.pref,sep="."))

  #print(filt.reflib.path)

  filt_txt=paste("input cutoff",i, sep=" ")


  #histograms
  #Reads per RSL-guide
  title_plot1=paste("Filtering: ",filt_txt,sep=(" "))

  freq_data=read.delim(file.path(filt.reflib.path,paste(refdata.pref,i,"frequencies.filt_reads.tsv",sep=("."))), header = TRUE, sep = "\t", quote = "\"", fill = TRUE, row.names=NULL)
  raw_freq_r = gather(freq_data, variable, value,-readcount)
  colnames(raw_freq_r)=c("readcount","sample","frequency")


  plot1=ggplot(data=raw_freq_r, aes(x=readcount, y=frequency, colour=sample, fill=sample)) + geom_bar(stat="identity") +
  facet_wrap(~sample)+
    scale_fill_viridis(discrete=TRUE,option="turbo")+
    scale_colour_viridis(discrete=TRUE,option="turbo")+
    theme_bw() + theme(plot.margin = unit(c(0.5, -0.5, 0, -0.5), "cm"))+ 
    labs(title=title_plot1, y="count",x="reads per RSL-guide")+
    theme(aspect.ratio = 2/(1+sqrt(5)) ) + theme(legend.position="none",legend.margin=unit(c(0, -0.5, 0, -0.5), "cm"))


  x_lim=50 
  freqs.lim.i=raw_freq_r%>%filter(readcount>=1, readcount<=x_lim)
  y_lim=round(max(freqs.lim.i$frequency)+0.1*(max(freqs.lim.i$frequency)))


  plot2=ggplot(data=raw_freq_r, aes(x=readcount, y=frequency, colour=sample, fill=sample)) + geom_bar(stat="identity") +
  facet_wrap(~sample)+
    scale_fill_viridis(discrete=TRUE,option="turbo")+
    scale_colour_viridis(discrete=TRUE,option="turbo")+
    theme_bw() + theme(plot.margin = unit(c(0.5, -0.5, 0, -0.5), "cm"))+ 
    coord_cartesian(xlim=c(0,x_lim),ylim=c(0,y_lim) )+
    labs(title=title_plot1, y="count",x="reads per RSL-guide")+
    theme(aspect.ratio = 2/(1+sqrt(5)) ) + theme(legend.position="none",legend.margin=unit(c(0, -0.5, 0, -0.5), "cm"))


  hist.reads_per_RSLsgRNA[[paste0("CO.",i)]] = plot1
  hist.reads_per_RSLsgRNA2[[paste0("CO.",i)]] = plot2


  ### histogram of reads / RSL-guide with x log scale

raw_freq_r_n0=raw_freq_r%>%filter(readcount>0)


  plot3=ggplot(data=raw_freq_r_n0, aes(x=readcount, y=frequency, colour=sample, fill=sample)) + geom_bar(stat="identity") +
  facet_wrap(~sample)+
    scale_fill_viridis(discrete=TRUE,option="turbo")+
    scale_colour_viridis(discrete=TRUE,option="turbo")+
    theme_bw() + theme(plot.margin = unit(c(0.5, -0.5, 0, -0.5), "cm"))+ 
    coord_cartesian(xlim=c(1,x_lim),ylim=c(0,y_lim) )+
    scale_x_continuous(trans='log10') +
    labs(title=title_plot1, y="count",x="reads per RSL-guide")+
    theme(aspect.ratio = 2/(1+sqrt(5)) ) + theme(legend.position="none",legend.margin=unit(c(0, -0.5, 0, -0.5), "cm"))


  hist.reads_per_RSLsgRNA2.log[[paste0("CO.",i)]] = plot3


  #RSLs per guide
  UMI_perguide_data=read.delim(file.path(filt.reflib.path,paste(refdata.pref,i,"RSL.perguide.tsv",sep=("."))), header = TRUE, sep = "\t", quote = "\"", fill = TRUE, row.names=NULL)
  UMI_perguide_dat = gather(UMI_perguide_data[,-2], variable, value,-sgRNA)
  colnames(UMI_perguide_dat)=c("guide","sample","count")

  title_hist=paste("Filtering: ",filt_txt,sep=(" "))


  hist_unlim=ggplot(UMI_perguide_dat, aes(x=count, fill=sample)) +
    geom_histogram(binwidth=1, position = 'identity') +
    scale_fill_viridis(discrete=TRUE,option="turbo")+
    labs(title="",x="RSL per guide",y="count") + theme_bw() +
    facet_wrap(~sample) + theme(aspect.ratio = 2/(1+sqrt(5)) )+
    labs(title=title_hist, y="count",x="RSL per guide") +
    theme(legend.position="none") + theme(plot.margin = unit(c(0.5, 0, 0, 0), "cm"))

    hist.RSL_per_sgRNA[[paste0("CO.",i)]]=hist_unlim


  #boxplots
  title_box=paste("Filtering: ",filt_txt,sep=(" "))

  box_UMI_perguide=ggplot(UMI_perguide_dat, aes(x=sample, y=count))  + geom_boxplot()+
    theme_bw()+ theme(plot.margin = unit(c(0.5, 0, 0, 0), "cm"))+ 
    theme(axis.text.x=element_text(angle=90,hjust=1,size = 10), axis.text.y = element_text(size=10)) +
      labs(title=title_box, y="RSL per guide",x="") +
      coord_flip()


  box_UMI_perguide_log=ggplot(UMI_perguide_dat, aes(x=sample, y=count))  + geom_boxplot()+
    theme_bw()+ scale_y_log10()+ theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) +
    theme(axis.text.x=element_text(angle=90,hjust=1,size = 20), axis.text.y = element_text(size=20)) +
      labs(title=title_box, y="Log10 of RSL per guide",x="") +
      coord_flip()

    box.RSL_per_sgRNA[[paste0("CO.",i)]]=box_UMI_perguide


  # RSL data
  RSL_data=read.delim(file.path(filt.reflib.path,paste(refdata.pref,i,"RSL.perguide.tsv",sep=("."))), header = TRUE, sep = "\t", quote = "\"", fill = TRUE, row.names=NULL)

  detected_sgRNAs[[paste0("CO.",i)]]=nrow(RSL_data)



  dat=RSL_data[,-2]
  rownames(dat)=dat$sgRNA
  dat=dat[,-1]
  dat=as.data.frame(t(dat))
  dat$sample=rownames(dat)
  df_to_plot=melt(dat,id.vars="sample")
  colnames(df_to_plot)=c("sample","guide","RSLcount")
  df_to_plot$filtering=paste0("CO.",i)

  RSL_data.list[[paste0("CO.",i)]]=df_to_plot

}


```




```{r fig.align="center", fig.width=8,fig.height=16, echo = FALSE}

plotheights=rep(8,ceiling(length(my.cutoffs)/2))

grid.arrange(grobs = hist.reads_per_RSLsgRNA, ncol=2, top = "Reads per RSL-sgRNA combination. Complete axes range.", heights=unit(plotheights,"cm"))

#grid.arrange(grobs = hist.reads_per_RSLsgRNA, ncol=2, top = "Reads per RSL-sgRNA combination. Complete axes range.",
#  set_panel_size( width = unit(7, "cm")))

```
<br>




## Histogram plots - limited data range

Reads per RSL-guide plotted with the reduced range on the X and Y axes displaying the lower-left quandrant of the plot 
(i.e. region where most of the signal can be observed, as most RSL-sgRNA combinations do not tend to gather high numbers of reads).



```{r fig.align="center", fig.width=8,fig.height=16, echo = FALSE}

plotheights=rep(8,ceiling(length(my.cutoffs)/2))

grid.arrange(grobs = hist.reads_per_RSLsgRNA2, ncol=2, top = "Reads per RSL-sgRNA combination. Reduced axes range.", heights=unit(plotheights,"cm"))


```

<br>
<br>




## Histogram plots - limited data range, X axis log10-transformed

Reads per RSL-guide plotted with the reduced range on both Y axis and X axis; X axis on log10 scale.




```{r fig.align="center", fig.width=8,fig.height=16, echo = FALSE}

plotheights=rep(8,ceiling(length(my.cutoffs)/2))

grid.arrange(grobs = hist.reads_per_RSLsgRNA2.log, ncol=2, top = "Reads per RSL-sgRNA combination. Reduced axes range. X axis log10 transformed.", heights=unit(plotheights,"cm"))


```

<br>
<br>





## Histogram plots - RSL per guide

Number of RSL per guide is presented on the following plots. Too high filtering cutoff results in more sgRNAs with low RSL count, which is manifested by signal flattening in the higher ranges of RSL per sgRNA (X axis).


```{r fig.align="center", fig.width=8,fig.height=16, echo = FALSE}
plotheights=rep(8,ceiling(length(my.cutoffs)/2))

grid.arrange(grobs = hist.RSL_per_sgRNA, ncol=2, top = "Number of distinct RSL per guide.", heights=unit(plotheights,"cm"))


```

<br>

```{r fig.align="center", fig.width=8,fig.height=10, echo = FALSE}
grid.arrange(grobs = box.RSL_per_sgRNA, ncol=2, top = "Number of distinct RSL per guide.")


```


`r {paste("## Number of detected guides in reference dataset", refdata.pref, sep=" ")}`



```{r refdata-stats-table, include=TRUE,cache=FALSE, results='asis'}

refdata.sgRNA.df <- data.frame(matrix(unlist(detected_sgRNAs), nrow=length(detected_sgRNAs), byrow=TRUE))
refdata.sgRNA.df$CO=my.cutoffs
refdata.sgRNA.df$CO=as.numeric(as.character(refdata.sgRNA.df$CO))
refdata.sgRNA.df=refdata.sgRNA.df[order(refdata.sgRNA.df$CO,decreasing=FALSE),]
refdata.sgRNA.df=refdata.sgRNA.df[,c(2,1)]


title.2=paste0("Number of detected sgRNA guides under different filtering cut-off settings in reference dataset ",refdata.pref,".")
footnote.2="Number of sgRNA detected in all samples of the reference dataset after the filtering of the input library using the indicated cutoff."

kable(refdata.sgRNA.df, align = c("l", "c", "c"), booktabs = TRUE, row.names = FALSE, caption = title.2, 
  col.names=c("Input filtering Cut-Off","Number of detected sgRNAs") )%>%
    footnote(
      general = footnote.2,
      general_title = "Note.",
      footnote_as_chunk = TRUE
      )%>%
  kable_minimal(full_width = TRUE)

```



## Relationships between data from different filtering strategies

Scatterplots are used to visualise RSL counts for corresponding guides under different filtering strategies.

Each point corresponds to one sgRNA. RSL counts under different cutoffs are compared to each other.

Data from sample 



`r colnames(RSL_data)[3]`






```{r include=TRUE,cache=FALSE,results='asis'}




df_to_plot_all=bind_rows(RSL_data.list)
df_to_plot_all_smpl=df_to_plot_all[df_to_plot_all$sample==colnames(RSL_data)[3],]
df_scatter=spread(df_to_plot_all_smpl,filtering,RSLcount)


my.cutoffs2=paste0("CO.",my.cutoffs)

cutoff.pairs.mtx=combn(my.cutoffs2,2)
cutoff.pairs.number=ncol(cutoff.pairs.mtx)


scatters=list()

for (i in c(1:cutoff.pairs.number)){

 cutoff.pair.i=cutoff.pairs.mtx[,i]


  scatter.i=ggplot(df_scatter, aes_string(x=cutoff.pair.i[1],y=cutoff.pair.i[2]))+ geom_point() + geom_abline() +geom_jitter(width = 0.1, height = 0.1) + theme_bw()

  scatters[[paste0("plot.",i)]] = scatter.i

}


```


<br>

```{r ig.align="center", fig.width=8, fig.height=32, echo = FALSE}
grid.arrange(grobs = scatters, ncol=2, top = "Scatter plots RSL counts per sgRNA under different filtering cut offs.")

```

