---
title: Comparison of Filtering Cut-Offs
author: CRISPR Functional Genomics
output:
  bookdown::html_document2:
    theme: journal
    highlight: kate
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
date: "`r format(Sys.time(), '%d %B, %Y')`"
---



Run by: `r Sys.getenv("USER")` 


 The rationale and process:

* Not all theoretical combinations of sgRNA - RSL barcodes are present in the input library; Our goal is to detect the combinations which are artefacts arisen due to sequencing errors; These artefacts, which typically are present in only one technical replicate and / or are detected by very small number read count give rise to the zero inflated signal in downstreal RSL analysis;

* Input library has been sequenced in two technical replicates to allow for detection of these artefactual reads;

* sgRNA - RSL combinations are retained if they are present in **both replicates** and have a **read count >= cutoff** in both replicates individually;

* Filtering cutoffs in the range of 0 (no filtering) to 10 (excessive filtering) are compared with respect to the following metrics:

  * number of retained sgRNA - RSL combinations in the filtered input library;

  * library sizes of the input library technical replicates after the filtering;


* Finally, each of the input libraries filtered using different cutoffs is used for processing a reference data set and the following metrics are evaluated:

  * number of detected sgRNA - RSL combinations in the reference data set after filtering;

  * number of detected sgRNAs in the reference data set after filtering;;

  * library sizes of the reference data set samples after filtering;

  * distribution of read counts per sgRNA - RSL combination (histogram plots);

  * distribution of RSL counts per sgRNA (histogram and box plots);

  * comparison of RSL counts per sgRNA between different strategies (scatter plots);



----------------

```{r, echo=FALSE}
```


```{r local-tsts, include=FALSE,cache=FALSE}
#path to count table for reference data
#refdata.path="/Users/agata.smialowska/NBISproj/5351_CRISPR_pipeline/data/ctables/Calle_Heldin.UMIcounts.csv"

#root dir for filtered input libraries
#inputdata.path="/Users/agata.smialowska/NBISproj/5351_CRISPR_pipeline/gh/rackham_tsts/input/Brunello_v2022/results/input_filtered"

#inputlib.name="Brunello_v2022"

#refdata.pref="Calle_Heldin.UMIcounts"


```


```{r include=FALSE,cache=FALSE}
library(knitr)
library(bookdown)


knitr::opts_chunk$set(echo = FALSE, 
                      collapse = TRUE, 
                      warning = FALSE, 
                      message = FALSE)

```




```{r dirs, include=FALSE,cache=FALSE}


```


```{r prep_environment, include=FALSE,cache=FALSE}
library(ggplot2)
library(tibble)
library(tidyr)
library(dplyr)
library(reshape2)
library(viridis)
library(parallel)
#library(cowplot)
library(gridExtra)

library(kableExtra)
```

# Analysis of Input Library `r inputlib.name`


```{r input-stats, include=FALSE,cache=FALSE}
fname_log=paste0(inputlib.name,".report.log")

resdirs=dir(inputdata.path)

#exclude dirs with refdata.pref
resdirs=resdirs[!grepl(refdata.pref, resdirs)]

#input.stats <- vector(mode = "list", length = length(resdirs))

input.stats <- vector(mode = "list")


for (i in resdirs){
  input.stats.i=file.path(inputdata.path,i,fname_log)
  print("Processing")
  print(input.stats.i)

  input.metrics.i=read.table(input.stats.i, sep="\t", header=FALSE, blank.lines.skip=TRUE)

  input.stats[[i]]=input.metrics.i

}


input.stats.df <- data.frame(matrix(unlist(input.stats), nrow=length(input.stats), byrow=TRUE))
colnames(input.stats.df)=c("CO","sgRNA-RSL","ColSums")
input.stats.df$CO=as.numeric(as.character(input.stats.df$CO))
input.stats.df=input.stats.df[order(input.stats.df$CO,decreasing=FALSE),]

```

```{r input-stats-table, include=TRUE,cache=FALSE, results='asis'}

title.1="Number of detected sgRNA-RSL combination under different filtering cut-off settings."

footnote.1="Number of sgRNA-RSL combinations and library sizes are calculated after the filtering of the input library using the indicated cutoff."

kable(input.stats.df, align = c("l", "c", "c"), booktabs = TRUE, row.names = FALSE, caption = title.1, 
  col.names=c("Cut Off","Number of sgRNA-RSL Combinations","Library Sizes (in technical replicates)") )%>%
    footnote(
      general = footnote.1,
      general_title = "Note.",
      footnote_as_chunk = TRUE
      )%>%
  kable_minimal(full_width = TRUE)

```

# Effect of Filtering on Reference Data `r refdata.pref`

## Read and RSL counts

Reads per RSL-sgRNA histograms


```{r include=TRUE,cache=FALSE}

#my.cutoffs=c(1,2,3,4,5,7,10)

my.cutoffs=c(0,1,2,3,4,5,7,10)


hist.reads_per_RSLsgRNA <- list()
hist.reads_per_RSLsgRNA2 <- list()

hist.RSL_per_sgRNA <- list()

box.RSL_per_sgRNA <- list()

detected_sgRNAs <- list()

RSL_data.list <- list()



for (i in my.cutoffs){
  filt.reflib.path=file.path(inputdata.path,paste(inputlib.name,i,refdata.pref,sep="."))

  #print(filt.reflib.path)

  filt_txt=paste("input cutoff",i, sep=" ")


  #histograms
  #Reads per RSL-guide
  title_plot1=paste("Filtering: ",filt_txt,sep=(" "))

  freq_data=read.delim(file.path(filt.reflib.path,paste(refdata.pref,i,"frequencies.filt_reads.tsv",sep=("."))), header = TRUE, sep = "\t", quote = "\"", fill = TRUE, row.names=NULL)
  raw_freq_r = gather(freq_data, variable, value,-readcount)
  colnames(raw_freq_r)=c("readcount","sample","frequency")


  plot1=ggplot(data=raw_freq_r, aes(x=readcount, y=frequency, colour=sample, fill=sample)) + geom_bar(stat="identity") +
  facet_wrap(~sample)+
    scale_fill_viridis(discrete=TRUE,option="turbo")+
    scale_colour_viridis(discrete=TRUE,option="turbo")+
    theme_bw() + theme(plot.margin = unit(c(0.5, -0.5, 0, -0.5), "cm"))+ 
    labs(title=title_plot1, y="count",x="reads per RSL-guide")+
    theme(aspect.ratio = 2/(1+sqrt(5)) ) + theme(legend.position="none",legend.margin=unit(c(0, -0.5, 0, -0.5), "cm"))


  x_lim=40 
  y_lim=200000

  plot2=ggplot(data=raw_freq_r, aes(x=readcount, y=frequency, colour=sample, fill=sample)) + geom_bar(stat="identity") +
  facet_wrap(~sample)+
    scale_fill_viridis(discrete=TRUE,option="turbo")+
    scale_colour_viridis(discrete=TRUE,option="turbo")+
    theme_bw() + theme(plot.margin = unit(c(0.5, -0.5, 0, -0.5), "cm"))+ 
    coord_cartesian(xlim=c(0,x_lim),ylim=c(0,y_lim) )+
    labs(title=title_plot1, y="count",x="reads per RSL-guide")+
    theme(aspect.ratio = 2/(1+sqrt(5)) ) + theme(legend.position="none",legend.margin=unit(c(0, -0.5, 0, -0.5), "cm"))


  hist.reads_per_RSLsgRNA[[paste0("CO.",i)]] = plot1
  hist.reads_per_RSLsgRNA2[[paste0("CO.",i)]] = plot2


  #RSLs per guide
  UMI_perguide_data=read.delim(file.path(filt.reflib.path,paste(refdata.pref,i,"RSL.perguide.tsv",sep=("."))), header = TRUE, sep = "\t", quote = "\"", fill = TRUE, row.names=NULL)
  UMI_perguide_dat = gather(UMI_perguide_data[,-2], variable, value,-sgRNA)
  colnames(UMI_perguide_dat)=c("guide","sample","count")

  title_hist=paste("Filtering: ",filt_txt,sep=(" "))


  hist_unlim=ggplot(UMI_perguide_dat, aes(x=count, fill=sample)) +
    geom_histogram(binwidth=1, position = 'identity') +
    scale_fill_viridis(discrete=TRUE,option="turbo")+
    labs(title="",x="RSL per guide",y="count") + theme_bw() +
    facet_wrap(~sample) + theme(aspect.ratio = 2/(1+sqrt(5)) )+
    labs(title=title_hist, y="count",x="UMIs per guide") +
    theme(legend.position="none") + theme(plot.margin = unit(c(0.5, 0, 0, 0), "cm"))

    hist.RSL_per_sgRNA[[paste0("CO.",i)]]=hist_unlim


  #boxplots
  title_box=paste("Filtering: ",filt_txt,sep=(" "))

  box_UMI_perguide=ggplot(UMI_perguide_dat, aes(x=sample, y=count))  + geom_boxplot()+
    theme_bw()+ theme(plot.margin = unit(c(0.5, 0, 0, 0), "cm"))+ 
    theme(axis.text.x=element_text(angle=90,hjust=1,size = 10), axis.text.y = element_text(size=10)) +
      labs(title=title_box, y="RSL per guide",x="") +
      coord_flip()


  box_UMI_perguide_log=ggplot(UMI_perguide_dat, aes(x=sample, y=count))  + geom_boxplot()+
    theme_bw()+ scale_y_log10()+ theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) +
    theme(axis.text.x=element_text(angle=90,hjust=1,size = 20), axis.text.y = element_text(size=20)) +
      labs(title=title_box, y="Log10 of RSL per guide",x="") +
      coord_flip()

    box.RSL_per_sgRNA[[paste0("CO.",i)]]=box_UMI_perguide


  # RSL data
  RSL_data=read.delim(file.path(filt.reflib.path,paste(refdata.pref,i,"RSL.perguide.tsv",sep=("."))), header = TRUE, sep = "\t", quote = "\"", fill = TRUE, row.names=NULL)

  detected_sgRNAs[[paste0("CO.",i)]]=nrow(RSL_data)



  dat=RSL_data[,-2]
  rownames(dat)=dat$sgRNA
  dat=dat[,-1]
  dat=as.data.frame(t(dat))
  dat$sample=rownames(dat)
  df_to_plot=melt(dat,id.vars="sample")
  colnames(df_to_plot)=c("sample","guide","RSLcount")
  df_to_plot$filtering=paste0("CO.",i)

  RSL_data.list[[paste0("CO.",i)]]=df_to_plot

}


```



```{r fig.align="center", fig.width=8,fig.height=16, echo = FALSE}

plotheights=rep(8,ceiling(length(my.cutoffs)/2))

grid.arrange(grobs = hist.reads_per_RSLsgRNA, ncol=2, top = "Reads per RSL-sgRNA combination. Complete axes range.", heights=unit(plotheights,"cm"))

#grid.arrange(grobs = hist.reads_per_RSLsgRNA, ncol=2, top = "Reads per RSL-sgRNA combination. Complete axes range.",
#  set_panel_size( width = unit(7, "cm")))

```
<br>




Reads per RSL-guide plotted with the reduced range on the X and Y axes displaying the most interesting region of the plot.


```{r fig.align="center", fig.width=8,fig.height=16, echo = FALSE}

plotheights=rep(8,ceiling(length(my.cutoffs)/2))

grid.arrange(grobs = hist.reads_per_RSLsgRNA2, ncol=2, top = "Reads per RSL-sgRNA combination. Reduced axes range.", heights=unit(plotheights,"cm"))


```

<br>
<br>


Number of RSL per guide


```{r fig.align="center", fig.width=8,fig.height=16, echo = FALSE}
plotheights=rep(8,ceiling(length(my.cutoffs)/2))

grid.arrange(grobs = hist.RSL_per_sgRNA, ncol=2, top = "Number of distinct RSL per guide.", heights=unit(plotheights,"cm"))


```

<br>

```{r ig.align="center", fig.width=8,fig.height=10, echo = FALSE}
grid.arrange(grobs = box.RSL_per_sgRNA, ncol=2, top = "Number of distinct RSL per guide.")


```


## Number of detected guides in reference dataset `r refdata.pref`



```{r refdata-stats-table, include=TRUE,cache=FALSE, results='asis'}

refdata.sgRNA.df <- data.frame(matrix(unlist(detected_sgRNAs), nrow=length(detected_sgRNAs), byrow=TRUE))
refdata.sgRNA.df$CO=my.cutoffs
refdata.sgRNA.df$CO=as.numeric(as.character(refdata.sgRNA.df$CO))
refdata.sgRNA.df=refdata.sgRNA.df[order(refdata.sgRNA.df$CO,decreasing=FALSE),]
refdata.sgRNA.df=refdata.sgRNA.df[,c(2,1)]


title.2=paste0("Number of detected sgRNA guides under different filtering cut-off settings in reference dataset ",refdata.pref,".")
footnote.2="Number of sgRNA detected in all samples of the reference dataset after the filtering of the input library using the indicated cutoff."

kable(refdata.sgRNA.df, align = c("l", "c", "c"), booktabs = TRUE, row.names = FALSE, caption = title.2, 
  col.names=c("Input filtering Cut-Off","Number of detected sgRNAs") )%>%
    footnote(
      general = footnote.2,
      general_title = "Note.",
      footnote_as_chunk = TRUE
      )%>%
  kable_minimal(full_width = TRUE)

```


## Relationships between data from different filtering strategies

Scatterplots are used to visualise RSL counts for corresponding guides under different filtering strategies.

Each point corresponds to one sgRNA. RSL counts under different cutoffs are compared to each other.

Data from sample `r colnames(RSL_data)[3]`


```{r include=TRUE,cache=FALSE,results='asis'}




df_to_plot_all=bind_rows(RSL_data.list)
df_to_plot_all_smpl=df_to_plot_all[df_to_plot_all$sample==colnames(RSL_data)[3],]
df_scatter=spread(df_to_plot_all_smpl,filtering,RSLcount)


my.cutoffs2=paste0("CO.",my.cutoffs)

cutoff.pairs.mtx=combn(my.cutoffs2,2)
cutoff.pairs.number=ncol(cutoff.pairs.mtx)


scatters=list()

for (i in c(1:cutoff.pairs.number)){

 cutoff.pair.i=cutoff.pairs.mtx[,i]


  scatter.i=ggplot(df_scatter, aes_string(x=cutoff.pair.i[1],y=cutoff.pair.i[2]))+ geom_point() + geom_abline() +geom_jitter(width = 0.1, height = 0.1) + theme_bw()

  scatters[[paste0("plot.",i)]] = scatter.i

}


```


<br>

```{r ig.align="center", fig.width=8, fig.height=32, echo = FALSE}
grid.arrange(grobs = scatters, ncol=2, top = "Scatter plots RSL counts per sgRNA under different filtering cut offs.")


```

I to juz wszystko drogie dzieci. Teraz pocalujcie misia w dupe.

